Selective electroporation of the viral envelope

### 2021-05-07 documents/pure_electroporation.tex
Testing analytic control solution

\usepackage{physics}
\usepackage{bm}
\newcommand{\bstar}[1]{ {#1}^{\bm*} }

@@

The question of optimizing a pulse is not novel, having been discussed by many groups in the past. Retelj use a series of

\subsection*{Optimal pulse}

@@

\subsection{Attempts to solve in s-space; convolution integral}

@@

An objective function with a constant integral control power constraint, $t_{end} = \int_0^t_{end} u(t)^2 dt$ was used, but this did not appear to yield meaningful results.

@@

@@

The coefficients have a very small value; even time has a very small value. This was believed to account for at least some of the

$$\dv[2]{x_V}{t} = 
\frac{R a_{1V}}{b_{1V}} \dv[2]{u}{t} 
+ \frac{R a_{2V}}{b_{1V}}  \dv{u}{t} 
+ \frac{R a_{3V}}{b_{1V}}  u
- \frac{b_{2V}}{b_{1V}} \dv{x_V}{t}
- \frac{b_{3V}}{b_{1V}} x_V $$

$$\dv[2]{x_H}{t} = 
\frac{R a_{1H}}{b_{1H}} \dv[2]{u}{t} 
+ \frac{R a_{2H}}{b_{1H}}  \dv{u}{t} 
+ \frac{R a_{3H}}{b_{1H}}  u
- \frac{b_{2H}}{b_{1H}} \dv{x_H}{t}
- \frac{b_{3H}}{b_{1H}} x_H $$

%x2  = (R*a1*u2 + R*a2*u1 + R*a3*u0 - b2*x1 - b3*x0)/b1
% https://math.stackexchange.com/questions/3435173/how-to-nondimensionalize-a-second-order-differential-equation

let $x_V^{\bm*} = \frac{x_V}{x_{VK}}$ and $t^{\bm*} = \frac{t}{t_{K}}$ where $_K$ denotes a nondimensionalization constant ($t_0$ is switched to $t_K$ to avoid confusion with x0 used in simulation). (the forcing functional u is dependent on t, but this is implied).\\ \\

$$\frac{d^n}{dt^n} = \frac{1}{{t_K}^n}\frac{d^n}{d\bstar{t}^n}$$\\

$$\dv[2]{x_V}{t} = 
\frac{R a_{1V}}{b_{1V}} \dv[2]{u}{t} 
+ \frac{R a_{2V}}{b_{1V}}  \dv{u}{t} 
+ \frac{R a_{3V}}{b_{1V}}  u
- \frac{b_{2V}}{b_{1V}} \dv{x_V}{t}
- \frac{b_{3V}}{b_{1V}} x_V $$

@@

denormalize u?

% Is the effect replicable, or is it simply an electrochemical artifact? Will these nanopores be destructive to viruses - will the pores be open for sufficient time for the ssRNA to be released? Will the nucleocapsid restrain the genome sufficiently and prevent membrane damage from inactivating the virus? Are the apoptosis effects likely to be triggered before these? Are there similar structures in tissue?

% A serious danger would be that if a nuclear or organelle membrane were to be porated, viral ssRNA may be admitted to places it shouldn't be. The author does not have the biology experience to evaluate this risk.

>>>>

\usepackage{physics}
\usepackage{bm}
\usepackage{listings}
\newcommand{\bstar}[1]{ {#1}^{\bm*} }

@@

The concept of optimizing an electroporation pulse is not novel, having been discussed by many groups in the past; either to prodduce different sterilization effects, tee Retelj use a series of

\subsection{Attempts to solve for the optimal pulse in the s-space; convolution integral}

@@

@@

An objective function with a constant integral control power constraint, $t_{end} = \int_0^{t_{end}} u(t)^2 dt$ was used, but this did not appear to yield meaningful results.

@@

iLQR. Differential dynamic programming was not evaluated (julia lib).

@@

The coefficients all have a very small value in standard units. Numerical error was presumed to account for at least some of the difficulty in optimization, so was partially nondimensionalized.

the a and b coefficients both have units

\[ \{\text{meter}\cdot \text{siemens}^2, \  \text{farad}\cdot \text{meter}\cdot \text{siemens} ,\ \text{farad}^2 \cdot \text{meter}\} \]

in physical units, typical values for as an example, \\
\verb|Cell(0.3, 80, 0.005, 30, 1e-8, 60, 50e-9, 14e-9, t)| \

$$R = 2.5\times10^{-8}$$
$$a_1 = 6\times10^{-26}$$
$$a_2 = 9\times10^{-33}$$
$$a_3 = 2\times10^{-41}$$
$$b_1 = 4\times10^{-26}$$
$$b_2 = 1\times10^{-32}$$
$$b_3 = 4\times10^{-41}$$

An input control waveform u(t) (volts per meter), and two output waveforms, $x_V(t)$ and $x_H(t)$ in volts.\\

%virus.__dict__

First system of differential equations (named V):\\

$$\dv[2]{x_V}{t} = 
\frac{R_V a_{1V}}{b_{1V}} \dv[2]{u}{t} 
+ \frac{R_V a_{2V}}{b_{1V}}  \dv{u}{t} 
+ \frac{R_V a_{3V}}{b_{1V}}  u
- \frac{b_{2V}}{b_{1V}} \dv{x_V}{t}
- \frac{b_{3V}}{b_{1V}} x_V $$

In practice, the solver takes a DAE: the second differential equation is split into an explicit system of eqs:

$$u1 = \dv{}{t} u$$
$$u2 = \dv{}{t} u1$$

$$x1_V = \dv{}{t} x0_V$$
$$x2_V = \dv{}{t} x1_V$$

Another system is solved simultaneously (named H), of identical form but different coefficients:

$$\dv[2]{x_H}{t} = 
\frac{R_H a_{1H}}{b_{1H}} \dv[2]{u}{t} 
+ \frac{R_H a_{2H}}{b_{1H}}  \dv{u}{t} 
+ \frac{R_H a_{3H}}{b_{1H}}  u
- \frac{b_{2H}}{b_{1H}} \dv{x_H}{t}
- \frac{b_{3H}}{b_{1H}} x_H $$

%x2  = (R*a1*u2 + R*a2*u1 + R*a3*u0 - b2*x1 - b3*x0)/b1

Let $x_V^{\bm*} = \frac{x_V}{x_{VK}}$ and $t^{\bm*} = \frac{t}{t_{K}}$ where $_K$ denotes a nondimensionalization constant (standard convention of $t_0$ is changed to $t_K$ to avoid confusion with subscript 0s used in simulation). (the forcing functional u is dependent on t, but this is implied).\\ \\

$$\frac{d^n}{dt^n} = \frac{1}{{t_K}^n}\frac{d^n}{d{\bstar{t}}^n}$$\\

% https://math.stackexchange.com/questions/3435173/how-to-nondimensionalize-a-second-order-differential-equation
% http://math.colgate.edu/~wweckesser/math312Spring05/handouts/Nondim.pdf
% https://faculty.math.illinois.edu/~rdeville/teaching/558/nondimensionalization.pdf
% http://math.colgate.edu/~wweckesser/math312Spring05/handouts/MoreNondim.pdf

% Kronbichler et al suggest an interesting method of nondimensionalizing.

% https://www.math.colostate.edu/~bangerth/publications/2008-boussinesq.pdf

@@

\subsection*{}

However, this overlooks a much simpler method.

Setting say, the logistic equation, produces an overdetermined system. using the above as a constraint in PSOPT was also ineffective.
